---
title: "Ireland County Sea-Distance Dashboard"
format:
  dashboard:
    orientation: pages
    scrolling: false
    theme: cosmo
execute:
  echo: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false

library(sf)
library(leaflet)
library(dplyr)
library(igraph)

# ---- helpers (inline so the vignette is self-contained) ----------------------

.get_ireland_32_counties <- function() {
  # ROI: 26 counties from GADM IRL level 1
  irl <- sf::st_read(
    "https://geodata.ucdavis.edu/gadm/gadm4.1/json/gadm41_IRL_1.json",
    quiet = TRUE
  )
  irl$county_name <- ifelse(
    irl$HASC_1 == "IE.CK", "Cork", as.character(irl$NAME_1)
  )
  irl$country <- "Republic of Ireland"
  irl <- irl[, c("county_name", "country", "geometry")]

  # NI: dissolve 26 pre-2015 district councils -> 6 traditional counties
  gbr3 <- sf::st_read(
    "https://geodata.ucdavis.edu/gadm/gadm4.1/json/gadm41_GBR_3.json",
    quiet = TRUE
  )
  ni3 <- gbr3[!is.na(gbr3$NAME_1) & gbr3$NAME_1 == "NorthernIreland", ]

  ni_lookup <- c(
    Antrim = "Antrim", Newtownabbey = "Antrim", Ballymena = "Antrim",
    Ballymoney = "Antrim", Carrickfergus = "Antrim", Larne = "Antrim",
    Moyle = "Antrim", Belfast = "Antrim",
    Armagh = "Armagh", Banbridge = "Armagh", Craigavon = "Armagh",
    Ards = "Down", NorthDown = "Down", Castlereagh = "Down",
    Down = "Down", Lisburn = "Down", NewryandMourne = "Down",
    Fermanagh = "Fermanagh",
    Coleraine = "Londonderry", Derry = "Londonderry",
    Limavady = "Londonderry", Magherafelt = "Londonderry",
    Cookstown = "Tyrone", Dungannon = "Tyrone", Omagh = "Tyrone",
    Strabane = "Tyrone"
  )

  ni3$county_name <- ni_lookup[ni3$NAME_3]
  ni3 <- ni3[!is.na(ni3$county_name), ]

  ni_counties <- stats::aggregate(
    ni3["geometry"],
    by = list(county_name = ni3$county_name),
    FUN = function(x) sf::st_union(x)
  )
  ni_counties$country <- "Northern Ireland"
  ni_counties <- ni_counties[, c("county_name", "country", "geometry")]

  rbind(irl, ni_counties)
}

.get_coastline <- function(year = 2021) {
  giscoR::gisco_get_coastallines(year = year, resolution = "10")
}

.compute_sea_distance <- function(counties, coastline) {
  adj_matrix <- sf::st_touches(counties, sparse = FALSE) * 1L
  g <- igraph::graph_from_adjacency_matrix(adj_matrix, mode = "undirected",
                                           diag = FALSE)
  coastal     <- lengths(sf::st_intersects(counties, coastline)) > 0
  coastal_idx <- which(coastal)
  n           <- nrow(counties)
  dist        <- rep(NA_integer_, n)

  if (length(coastal_idx) > 0 && length(coastal_idx) < n) {
    sp   <- igraph::distances(g, v = coastal_idx)
    dist <- as.integer(apply(sp, 2, min))
  } else if (length(coastal_idx) == n) {
    dist <- rep(0L, n)
  }
  counties$sea_distance <- dist
  counties
}

.summarise_sea_distance <- function(counties) {
  dplyr::count(sf::st_drop_geometry(counties), .data$sea_distance, name = "n")
}

# ---- fetch data ---------------------------------------------------------------
counties  <- .get_ireland_32_counties()
coastline <- .get_coastline()
counties  <- .compute_sea_distance(counties, coastline)
summary_df <- .summarise_sea_distance(counties)

# ---- derived values for dynamic prose ----------------------------------------
n_counties    <- nrow(counties)
n_coastal     <- sum(counties$sea_distance == 0L, na.rm = TRUE)
n_one_hop     <- sum(counties$sea_distance == 1L, na.rm = TRUE)
n_two_hop     <- sum(counties$sea_distance == 2L, na.rm = TRUE)
max_distance  <- max(counties$sea_distance, na.rm = TRUE)
pct_coastal   <- round(100 * n_coastal / n_counties, 1)

coastal_names <- sort(counties$county_name[counties$sea_distance == 0L])
inland_names  <- sort(counties$county_name[counties$sea_distance == max_distance])

# ---- colour palette (viridis, colorblind-safe) --------------------------------
all_levels    <- as.character(seq(0L, max_distance))
viridis_ramp  <- scales::viridis_pal(option = "D")(length(all_levels))
names(viridis_ramp) <- all_levels

pal_fn <- leaflet::colorFactor(
  palette = viridis_ramp,
  domain  = as.character(counties$sea_distance)
)

counties_wgs84 <- sf::st_transform(counties, crs = 4326)
```

# Ireland Map {orientation="columns"}

## Column {width="70%"}

```{r}
#| label: fig-map
#| fig-cap: !expr paste0(
#|   "Interactive choropleth of Ireland's ", n_counties, " traditional counties coloured by sea-distance. ",
#|   "Sea-distance is the minimum number of shared-boundary county-hops to reach the coast ",
#|   "(distance 0 = directly coastal, 1 = one hop inland, etc.). ",
#|   "Colour scale uses the viridis palette (purple to green) for colorblind accessibility. ",
#|   "Of the ", n_counties, " counties, ", n_coastal, " (", pct_coastal, "%) touch the coast directly. ",
#|   "The most inland counties (distance ", max_distance, ") are: ",
#|   paste(inland_names, collapse = ", "), ". ",
#|   "Click any county to see its name, jurisdiction (ROI/NI), and sea-distance value. ",
#|   "Source: GADM 4.1 (26 ROI counties + 6 NI traditional counties via dissolve); ",
#|   "coastline from GISCO 2021."
#| )

leaflet(data = counties_wgs84) |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addPolygons(
    fillColor   = ~ pal_fn(as.character(sea_distance)),
    fillOpacity = 0.75,
    color       = "#ffffff",
    weight      = 1,
    opacity     = 1,
    popup       = ~ paste0(
      "<strong>", county_name, "</strong><br>",
      country, "<br>",
      "Sea distance: ", sea_distance
    ),
    label       = ~ paste0(county_name, " (", sea_distance, ")")
  ) |>
  leaflet::addLegend(
    position = "bottomright",
    pal      = pal_fn,
    values   = ~ as.character(sea_distance),
    title    = "Sea distance<br>(county hops)",
    opacity  = 0.85
  )
```

## Column {width="30%"}

### Key figures

```{r}
#| label: tbl-key-stats

knitr::kable(
  data.frame(
    Metric = c(
      "Total counties",
      "Coastal (distance 0)",
      "One hop inland (distance 1)",
      paste0(max_distance, " hops inland (distance ", max_distance, ")"),
      "Maximum sea distance",
      "% directly coastal"
    ),
    Value = c(
      n_counties,
      n_coastal,
      n_one_hop,
      n_two_hop,
      max_distance,
      paste0(pct_coastal, "%")
    )
  ),
  align = c("l", "r"),
  col.names = c("Metric", "Value"),
  caption = paste0(
    "Summary statistics for Ireland's ", n_counties, " traditional counties. ",
    "Sea distance is computed via breadth-first search on the county-adjacency graph. ",
    "Coastal counties (those intersecting the GISCO coastline geometry) have distance 0; ",
    "each subsequent hop adds 1. ",
    pct_coastal, "% of counties touch the coast directly. ",
    "Source: GADM 4.1 (ROI + NI); coastline from GISCO 2021."
  )
)
```

### About the sea-distance metric

Ireland has `r n_counties` traditional counties (`r n_counties - 6L` in the
Republic, 6 in Northern Ireland). Sea distance measures how many
county-boundaries separate a county from the nearest coastal county.

Of these, `r n_coastal` counties (`r pct_coastal`%) touch the coast directly.
`r n_one_hop` county(ies) are one hop from the coast, and
`r n_two_hop` county(ies) are `r max_distance` hops inland.

The most inland counties are: **`r paste(inland_names, collapse = ", ")`**.
The maximum sea distance is **`r max_distance`**, reflecting Ireland's
compact geography relative to its extensive coastline.

---

*Data: 26 ROI counties from [GADM 4.1](https://gadm.org/), 6 NI counties
dissolved from GADM district councils. Coastline from
[GISCO 2021](https://ec.europa.eu/eurostat/web/gisco). Adjacency via
`sf::st_touches()`; BFS via [igraph](https://r.igraph.org/).*

# Summary {orientation="columns"}

## Column {width="50%"}

```{r}
#| label: tbl-summary

knitr::kable(
  summary_df |>
    dplyr::mutate(
      pct = round(100 * n / sum(n), 1),
      label = dplyr::case_when(
        sea_distance == 0L ~ "Coastal (0 hops)",
        sea_distance == 1L ~ "One hop inland",
        TRUE               ~ paste0(sea_distance, " hops inland")
      )
    ) |>
    dplyr::select(label, n, pct) |>
    dplyr::rename(
      "Sea distance"       = label,
      "Count"              = n,
      "Share (%)"          = pct
    ),
  align   = c("l", "r", "r"),
  caption = paste0(
    "Distribution of Ireland's ", n_counties, " counties across sea-distance categories. ",
    "Each row corresponds to one distance level (0 = coastal, 1 = one shared-boundary hop, etc.). ",
    "'Share' is the percentage of all ", n_counties, " counties at that distance. ",
    "Coastal and near-coastal counties dominate: ",
    n_coastal + n_one_hop, " out of ", n_counties, " lie within one hop of the sea (",
    round(100 * (n_coastal + n_one_hop) / n_counties, 1), "%). ",
    "Source: GADM 4.1 (ROI + NI); sea distance via igraph BFS."
  )
)
```

## Column {width="50%"}

```{r}
#| label: fig-bar
#| fig-cap: !expr paste0(
#|   "Bar chart showing the count of Ireland's ", n_counties, " traditional counties ",
#|   "at each sea-distance level. ",
#|   "The x-axis gives sea distance in county hops (0 = directly coastal); ",
#|   "the y-axis gives the number of counties at that distance. ",
#|   "Bars use the same viridis palette as the map for visual correspondence. ",
#|   n_coastal, " counties are coastal and counts fall with increasing inland distance. ",
#|   "No county is more than ", max_distance, " hops from the sea. ",
#|   "Source: GADM 4.1 (ROI + NI); sea distance from igraph BFS."
#| )

barplot(
  height    = summary_df$n,
  names.arg = paste0(summary_df$sea_distance, "\nhop",
                     ifelse(summary_df$sea_distance == 1L, "", "s")),
  col       = viridis_ramp[as.character(summary_df$sea_distance)],
  border    = NA,
  xlab      = "Sea distance (county hops)",
  ylab      = "Number of counties",
  main      = paste0("Sea-Distance Distribution (N = ", n_counties, " counties)"),
  las       = 1
)
```
