---
title: "County Sea-Distance"
format:
  dashboard:
    orientation: rows
    scrolling: true
    theme:
      - cosmo
      - custom.scss
execute:
  echo: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false

library(sf)
library(leaflet)
library(dplyr)
library(igraph)
library(DT)

# ---- helpers (inline so the vignette is self-contained) ----------------------

# Local cache fallback: resolve RDS path for installed package or dev tree.
# system.file() returns "" when the package is not installed; find.package()
# locates the source tree when loaded via pkgload::load_all() in dev mode.
.cached_rds <- function(filename) {
  installed <- system.file("extdata", filename, package = "maps")
  if (nzchar(installed)) return(readRDS(installed))
  dev_path <- tryCatch(
    file.path(find.package("maps"), "inst", "extdata", filename),
    error = function(e) ""
  )
  if (nzchar(dev_path) && file.exists(dev_path)) {
    return(readRDS(dev_path))
  }
  # Last resort: path relative to vignette source file location
  vignette_dir <- tryCatch(dirname(normalizePath(knitr::current_input())), error = function(e) "")
  if (nzchar(vignette_dir)) {
    src_path <- normalizePath(
      file.path(vignette_dir, "..", "..", "inst", "extdata", filename),
      mustWork = FALSE
    )
    if (file.exists(src_path)) return(readRDS(src_path))
  }
  NULL
}

.get_ireland_32_counties <- function() {
  irl <- tryCatch(
    sf::st_read(
      "https://geodata.ucdavis.edu/gadm/gadm4.1/json/gadm41_IRL_1.json",
      quiet = TRUE
    ),
    error = function(e) NULL
  ) %||% .cached_rds("gadm41_IRL_1.rds")
  irl$county_name <- ifelse(
    irl$HASC_1 == "IE.CK", "Cork", as.character(irl$NAME_1)
  )
  irl$country <- "Republic of Ireland"
  irl <- irl[, c("county_name", "country", "geometry")]

  gbr3 <- tryCatch(
    sf::st_read(
      "https://geodata.ucdavis.edu/gadm/gadm4.1/json/gadm41_GBR_3.json",
      quiet = TRUE
    ),
    error = function(e) NULL
  ) %||% .cached_rds("gadm41_GBR_3.rds")
  ni3 <- gbr3[!is.na(gbr3$NAME_1) & gbr3$NAME_1 == "NorthernIreland", ]

  ni_lookup <- c(
    Antrim = "Antrim", Newtownabbey = "Antrim", Ballymena = "Antrim",
    Ballymoney = "Antrim", Carrickfergus = "Antrim", Larne = "Antrim",
    Moyle = "Antrim", Belfast = "Antrim",
    Armagh = "Armagh", Banbridge = "Armagh", Craigavon = "Armagh",
    Ards = "Down", NorthDown = "Down", Castlereagh = "Down",
    Down = "Down", Lisburn = "Down", NewryandMourne = "Down",
    Fermanagh = "Fermanagh",
    Coleraine = "Londonderry", Derry = "Londonderry",
    Limavady = "Londonderry", Magherafelt = "Londonderry",
    Cookstown = "Tyrone", Dungannon = "Tyrone", Omagh = "Tyrone",
    Strabane = "Tyrone"
  )

  ni3$county_name <- ni_lookup[ni3$NAME_3]
  ni3 <- ni3[!is.na(ni3$county_name), ]

  ni_counties <- stats::aggregate(
    ni3["geometry"],
    by = list(county_name = ni3$county_name),
    FUN = function(x) sf::st_union(x)
  )
  ni_counties$country <- "Northern Ireland"
  ni_counties <- ni_counties[, c("county_name", "country", "geometry")]

  rbind(irl, ni_counties)
}

.compute_sea_distance <- function(counties, coastline, crs_projected = 27700L) {
  old_s2 <- sf::sf_use_s2()
  on.exit(sf::sf_use_s2(old_s2), add = TRUE)
  sf::sf_use_s2(FALSE)

  has_crs <- !is.na(sf::st_crs(counties))
  if (has_crs) {
    counties_m   <- sf::st_transform(counties, crs_projected)
    counties_buf <- sf::st_buffer(counties_m, 200)
    nb_mat       <- sf::st_intersects(counties_buf, counties_m, sparse = FALSE)
    diag(nb_mat) <- FALSE
    adj_matrix   <- nb_mat * 1L
  } else {
    adj_matrix <- sf::st_touches(counties, sparse = FALSE) * 1L
  }

  g <- igraph::graph_from_adjacency_matrix(adj_matrix, mode = "undirected",
                                           diag = FALSE)

  geom_types <- unique(as.character(sf::st_geometry_type(sf::st_geometry(coastline))))
  if (all(geom_types %in% c("POLYGON", "MULTIPOLYGON"))) {
    bbox_buf      <- sf::st_bbox(counties) + c(-1, -1, 1, 1)
    coast_cropped <- sf::st_crop(coastline, bbox_buf)
    coast_lines   <- sf::st_boundary(sf::st_geometry(coast_cropped))
    coastline_use <- sf::st_union(coast_lines)
  } else {
    coastline_use <- sf::st_union(sf::st_geometry(coastline))
  }

  if (has_crs && !is.na(sf::st_crs(coastline_use))) {
    coastline_use <- sf::st_transform(coastline_use, sf::st_crs(counties))
  }

  coastal     <- lengths(sf::st_intersects(counties, coastline_use)) > 0
  coastal_idx <- which(coastal)

  n         <- nrow(counties)
  min_dist  <- rep(Inf, n)

  if (length(coastal_idx) > 0 && length(coastal_idx) < n) {
    sp       <- igraph::distances(g, v = coastal_idx)
    min_dist <- apply(sp, 2, min)
  } else if (length(coastal_idx) == n) {
    min_dist <- rep(0, n)
  }

  # Island fix: counties that are coastal but appear disconnected get Inf
  # from the BFS (island counties have no land neighbours).
  # Reset any Inf that belongs to a directly-coastal county to 0.
  min_dist[is.infinite(min_dist) & coastal] <- 0
  dist <- as.integer(min_dist)

  counties$sea_distance <- dist
  counties
}

.compute_neighbour_table <- function(counties, crs_projected = 27700L) {
  old_s2 <- sf::sf_use_s2()
  on.exit(sf::sf_use_s2(old_s2), add = TRUE)
  sf::sf_use_s2(FALSE)

  counties_m <- sf::st_transform(counties, crs_projected)
  n <- nrow(counties_m)

  counties_buf <- sf::st_buffer(counties_m, 200)
  nb_mat <- sf::st_intersects(counties_buf, counties_m, sparse = FALSE)
  diag(nb_mat) <- FALSE

  results <- lapply(seq_len(n), function(i) {
    nb_idx   <- which(nb_mat[i, ])
    nb_names <- counties_m$county_name[nb_idx]

    cb_i     <- sf::st_boundary(sf::st_geometry(counties_m)[i])
    cb_i_buf <- sf::st_buffer(cb_i, 100)

    total_km <- sum(vapply(nb_idx, function(j) {
      cb_j   <- sf::st_boundary(sf::st_geometry(counties_m)[j])
      inters <- tryCatch(sf::st_intersection(cb_i_buf, cb_j), error = function(e) NULL)
      if (is.null(inters) || all(sf::st_is_empty(inters))) return(0)
      sum(as.numeric(sf::st_length(inters)), na.rm = TRUE) / 1000
    }, numeric(1L)))

    data.frame(
      county_name        = counties_m$county_name[i],
      n_neighbours       = length(nb_idx),
      neighbour_names    = paste(sort(nb_names), collapse = ", "),
      boundary_length_km = round(total_km, 1)
    )
  })

  out <- do.call(rbind, results)
  out[order(-out$n_neighbours, -out$boundary_length_km), ]
}

.summarise_sea_distance <- function(counties) {
  dplyr::count(sf::st_drop_geometry(counties), .data$sea_distance, name = "n")
}

# ---- render helpers ----------------------------------------------------------

.build_palette <- function(max_dist) {
  pal <- c("0" = "#fee090", "1" = "#fc8d59", "2" = "#d73027")
  if (max_dist > 2L) {
    extra <- scales::viridis_pal(option = "C")(max_dist - 2L)
    for (d in seq(3L, max_dist)) pal[as.character(d)] <- extra[d - 2L]
  }
  pal
}

.render_map <- function(counties_wgs84, pal_fn) {
  leaflet(data = counties_wgs84, height = "400px") |>
    leaflet::addProviderTiles(leaflet::providers$CartoDB.DarkMatter) |>
    leaflet::addPolygons(
      fillColor   = ~ pal_fn(sea_distance),
      fillOpacity = 0.8,
      color       = "#333333",
      weight      = 1.5,
      opacity     = 1,
      popup       = ~ paste0(
        "<div style=\"font-size: 16px;\">",
        "<strong>", county_name, "</strong><br>",
        country, "<br>",
        "Sea distance: ", sea_distance,
        "</div>"
      ),
      label       = ~ paste0(county_name, " (", sea_distance, ")")
    ) |>
    leaflet::addLegend(
      position  = "topleft",
      pal       = pal_fn,
      values    = ~ sea_distance,
      title     = "Sea distance",
      opacity   = 1
    )
}

.render_key_stats <- function(r) {
  stats_df <- data.frame(
    Metric = c(
      "Total counties",
      "Coastal (distance 0)",
      paste0("Distance ", seq_len(r$max_distance),
             " (", seq_len(r$max_distance), " hop",
             ifelse(seq_len(r$max_distance) == 1L, "", "s"), ")"),
      "Maximum sea distance",
      "% directly coastal"
    ),
    Value = c(
      as.character(r$n_counties),
      as.character(r$dist_counts["d0"]),
      vapply(seq_len(r$max_distance),
             function(d) as.character(r$dist_counts[paste0("d", d)]),
             character(1L)),
      as.character(r$max_distance),
      paste0(r$pct_coastal, "%")
    )
  )
  knitr::kable(
    stats_df,
    align     = c("l", "r"),
    col.names = c("Metric", "Value"),
    caption   = paste0(
      "Summary statistics for ", r$label_long, ". ",
      r$n_coastal, " coastal, ", r$n_inland, " inland; ",
      paste(r$deepest_names, collapse = " and "),
      " is the most inland (distance ", r$max_distance, "). ",
      "Source: ", r$gadm_link, " + ", r$gisco_link, "."
    )
  )
}

.render_summary_table <- function(r) {
  df <- r$summary |>
    dplyr::mutate(
      pct   = round(100 * n / sum(n), 1),
      label = dplyr::case_when(
        sea_distance == 0L ~ "Coastal (0 hops)",
        sea_distance == 1L ~ "1 hop inland",
        TRUE               ~ paste0(sea_distance, " hops inland")
      )
    ) |>
    dplyr::select(label, n, pct)
  DT::datatable(
    df,
    colnames  = c("Sea distance", "Count", "Share (%)"),
    rownames  = FALSE,
    options   = list(dom = "t", paging = FALSE, ordering = TRUE),
    caption   = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; color: #ffffff;",
      paste0(
        "Distribution of ", r$label_long, ". ",
        r$n_coastal, " coastal, ", r$n_inland, " inland. ",
        paste(r$deepest_names, collapse = " and "),
        " is the most inland (distance ", r$max_distance, "). ",
        "Source: ", r$gadm_link, " + ", r$gisco_link,
        "; sea distance via ", r$igraph_link, " BFS."
      )
    )
  )
}

.render_dot_plot <- function(r) {
  par(mar = c(4, 8, 2, 1), bg = "black", fg = "white",
      col.axis = "white", col.lab = "white", col.main = "white")
  dotchart(
    x      = rev(r$summary$n),
    labels = rev(paste0("Distance ", r$summary$sea_distance)),
    pch    = 19,
    pt.cex = 2.5,
    color  = rev(r$dist_palette[as.character(r$summary$sea_distance)]),
    xlab   = paste0("Number of ", r$unit_label, "s"),
    main   = paste0("Sea-Distance Distribution (N = ", r$n_counties, ")"),
    lcolor = "grey40"
  )
  grid(col = "grey30", lty = 1)
}

.render_neighbour_table <- function(r) {
  df <- r$nb_table |>
    dplyr::select(county_name, n_neighbours, boundary_length_km, neighbour_names)
  DT::datatable(
    df,
    colnames  = c("County", "Neighbours", "Boundary (km)", "Neighbour names"),
    rownames  = FALSE,
    filter    = "top",
    options   = list(
      pageLength = 15,
      scrollX    = TRUE,
      order      = list(list(1, "desc")),
      columnDefs = list(
        list(width = "100px", targets = 0),
        list(width = "60px", targets = 1),
        list(width = "80px", targets = 2)
      )
    ),
    caption   = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; color: #ffffff;",
      paste0(
        "All ", r$n_counties, " ", r$label_long,
        " sorted by neighbours (click headers to re-sort). ",
        "Use filters to search. ",
        "Boundary lengths in EPSG:", r$crs_projected, ". ",
        "Source: ", r$gadm_link, "; adjacency via ", r$sf_link, "."
      )
    )
  )
}

# ---- source links for captions -----------------------------------------------
gadm_link   <- '<a href="https://gadm.org/data.html" target="_blank">GADM 4.1</a>'
gisco_link  <- '<a href="https://ec.europa.eu/eurostat/web/gisco/geodata/statistical-units" target="_blank">GISCO 2016</a>'
igraph_link <- '<a href="https://r.igraph.org/" target="_blank">igraph</a>'
sf_link     <- '<a href="https://r-spatial.github.io/sf/" target="_blank">sf</a>'

# ---- fetch base data ---------------------------------------------------------
coastline <- giscoR::gisco_get_coastallines(year = 2016, resolution = "03") %||%
  .cached_rds("coastline_03_ireland_uk.rds")

ireland_counties_raw <- .get_ireland_32_counties()

gbr2 <- tryCatch(
  sf::st_read(
    "https://geodata.ucdavis.edu/gadm/gadm4.1/json/gadm41_GBR_2.json",
    quiet = TRUE
  ),
  error = function(e) NULL
) %||% .cached_rds("gadm41_GBR_2.rds")
england_raw  <- gbr2[gbr2$NAME_1 == "England",  ]
scotland_raw <- gbr2[gbr2$NAME_1 == "Scotland", ]
wales_raw    <- gbr2[gbr2$NAME_1 == "Wales",    ]

for (obj_name in c("england_raw", "scotland_raw", "wales_raw")) {
  obj <- get(obj_name)
  obj$county_name <- as.character(obj$NAME_2)
  obj$country     <- as.character(obj$NAME_1)

  # Fix literal "NA" county names via HASC_2 lookup and clean camel-case
  # (GADM 4.1 GBR level 2 has 67 rows where NAME_2 is the string "NA")
  is_na_str <- is.na(obj$county_name) | obj$county_name == "NA"
  if (any(is_na_str)) {
    hasc_lookup <- c(
      "GB.TM" = "Tameside",       "GB.TD" = "Trafford",
      "GB.WZ" = "Walsall",        "GB.WT" = "Warrington",
      "GB.WR" = "Warwick",        "GB.WN" = "Wigan",
      "GB.WL" = "Wiltshire",      "GB.WA" = "Windsor and Maidenhead",
      "GB.WO" = "Wokingham",      "GB.WC" = "Worcestershire",
      "GB.BS" = "Bristol",        "GB.BU" = "Buckinghamshire",
      "GB.BR" = "Bury",           "GB.CM" = "Cambridgeshire",
      "GB.CB" = "Central Bedfordshire",
      "GB.CQ" = "Cheshire East",  "GB.BX" = "Barnsley",
      "GB.CO" = "Cornwall",       "GB.CT" = "Coventry",
      "GB.DB" = "Derbyshire",     "GB.DO" = "Devon",
      "GB.DC" = "Doncaster",      "GB.DS" = "Dorset",
      "GB.DY" = "Dudley",         "GB.EY" = "East Riding of Yorkshire",
      "GB.ES" = "East Sussex",    "GB.EX" = "Essex",
      "GB.GL" = "Greater London", "GB.HA" = "Hampshire",
      "GB.HT" = "Hertfordshire",  "GB.KN" = "Knowsley",
      "GB.BI" = "Birmingham",     "GB.LC" = "Leicester",
      "GB.LE" = "Leicestershire", "GB.LI" = "Lincolnshire",
      "GB.MN" = "Manchester",     "GB.MB" = "Middlesbrough",
      "GB.NU" = "Newcastle upon Tyne",
      "GB.NF" = "Norfolk",        "GB.NL" = "North Lincolnshire",
      "GB.NS" = "North Somerset", "GB.NI" = "North Tyneside",
      "GB.NA" = "Northamptonshire", "GB.NB" = "Northumberland",
      "GB.NG" = "Nottingham",     "GB.NT" = "Nottinghamshire",
      "GB.OX" = "Oxfordshire",    "GB.PS" = "Portsmouth",
      "GB.RG" = "Reading",        "GB.RC" = "Redcar and Cleveland",
      "GB.RD" = "Rochdale",       "GB.RH" = "Rotherham",
      "GB.RL" = "Rutland",        "GB.BT" = "Bolton",
      "GB.ZF" = "Salford",        "GB.ZW" = "Sandwell",
      "GB.SE" = "Sefton",         "GB.SP" = "Shropshire",
      "GB.ZL" = "Slough",         "GB.SI" = "Solihull",
      "GB.SX" = "South Tyneside", "GB.ZH" = "Southampton",
      "GB.ST" = "Staffordshire",  "GB.SK" = "Stockport",
      "GB.SO" = "Stoke-on-Trent", "GB.SD" = "Sunderland",
      # Scotland
      "GB.NR" = "North Ayrshire",
      # Wales
      "GB.GD" = "Gwynedd",          "GB.MT" = "Merthyr Tydfil",
      "GB.NP" = "Newport",          "GB.SW" = "Swansea",
      "GB.TF" = "Torfaen",          "GB.VG" = "Vale of Glamorgan",
      "GB.WX" = "Wrexham",          "GB.BG" = "Bridgend",
      "GB.BJ" = "Blaenau Gwent",    "GB.DI" = "Denbighshire"
    )
    obj$county_name[is_na_str] <- hasc_lookup[obj$HASC_2[is_na_str]]
  }

  # Clean camel-case: insert space before capital following a lower-case letter
  obj$county_name <- gsub("([a-z])([A-Z])", "\\1 \\2", obj$county_name)
  # Remove trailing ", County of" / ", City of" suffixes
  obj$county_name <- gsub(",\\s*County\\s*of$|,Countyof$", "", obj$county_name)
  obj$county_name <- gsub(",\\s*City\\s*of$|,Cityof$",     "", obj$county_name)
  # Fix truncated Christchurch and Poole name
  obj$county_name <- gsub(",Christchurchand Po$", ", Christchurch and Poole", obj$county_name)

  # Drop rows where county_name is still NA after lookup (junk GADM rows)
  still_bad <- is.na(obj$county_name) | obj$county_name == "NA"
  if (any(still_bad)) {
    message("Dropping ", sum(still_bad), " rows with unresolvable county names in ", obj_name)
    obj <- obj[!still_bad, ]
  }

  # Validate: no NA or literal "NA" county names remain
  stopifnot(
    "county_name contains NA after name fixing"          = !any(is.na(obj$county_name)),
    "county_name contains literal 'NA' after name fixing" = !any(obj$county_name == "NA", na.rm = TRUE)
  )

  assign(obj_name, obj[, c("county_name", "country", "geometry")])
}

# ---- precompute all regions --------------------------------------------------
regions <- list()

region_specs <- list(
  Ireland  = list(raw = ireland_counties_raw,  crs = 2157L,  label_long = "Ireland's 32 traditional counties",  unit_label = "county"),
  England  = list(raw = england_raw,           crs = 27700L, label_long = "England's counties",                 unit_label = "county"),
  Scotland = list(raw = scotland_raw,          crs = 27700L, label_long = "Scotland's council areas",           unit_label = "council area"),
  Wales    = list(raw = wales_raw,             crs = 27700L, label_long = "Wales's counties",                   unit_label = "county")
)

for (rname in names(region_specs)) {
  spec      <- region_specs[[rname]]
  counties  <- .compute_sea_distance(spec$raw, coastline, crs_projected = spec$crs)
  stopifnot(!any(is.na(counties$sea_distance)))
  stopifnot(!any(is.na(counties$county_name)))

  summary_df  <- .summarise_sea_distance(counties)
  nb_table    <- .compute_neighbour_table(counties, crs_projected = spec$crs)

  n_counties   <- nrow(counties)
  n_coastal    <- sum(counties$sea_distance == 0L, na.rm = TRUE)
  n_inland     <- n_counties - n_coastal
  max_distance <- max(counties$sea_distance, na.rm = TRUE)
  pct_coastal  <- round(100 * n_coastal / n_counties, 1)
  deepest_names <- sort(counties$county_name[counties$sea_distance == max_distance])

  dist_counts <- setNames(
    vapply(0:max_distance,
           function(d) sum(counties$sea_distance == d, na.rm = TRUE),
           integer(1L)),
    paste0("d", 0:max_distance)
  )

  dist_palette <- .build_palette(max_distance)

  pal_fn <- leaflet::colorFactor(
    palette = unname(dist_palette[as.character(0:max_distance)]),
    domain  = 0:max_distance
  )

  counties_wgs84 <- sf::st_transform(counties, crs = 4326L)

  cap_map <- paste0(
    "Interactive choropleth of ", spec$label_long, " coloured by sea-distance. ",
    "Sea-distance is the minimum number of shared-boundary hops to reach the coast ",
    "(distance 0 = directly coastal, 1 = one hop inland, etc.). ",
    "Of the ", n_counties, " ", spec$unit_label, "s, ",
    n_coastal, " (", pct_coastal, "%) touch the coast directly ",
    "and ", n_inland, " are inland. ",
    "The most inland (distance ", max_distance, ") is ",
    paste(deepest_names, collapse = " and "), ". ",
    "Click any area for details. ",
    "Source: ", gadm_link, " + ", gisco_link, " coastline."
  )

  cap_dot <- paste0(
    "Dot plot showing the count of ", spec$label_long, " at each sea-distance level. ",
    "The x-axis gives the number of ", spec$unit_label, "s; ",
    "the y-axis gives sea distance in hops (0 = directly coastal). ",
    n_coastal, " are coastal (distance 0); ",
    "the most inland (distance ", max_distance, ") is ",
    paste(deepest_names, collapse = " and "), ". ",
    "Source: ", gadm_link, "; sea distance via ", igraph_link, " BFS."
  )

  regions[[rname]] <- list(
    counties      = counties,
    counties_wgs84 = counties_wgs84,
    summary       = summary_df,
    nb_table      = nb_table,
    n_counties    = n_counties,
    n_coastal     = n_coastal,
    n_inland      = n_inland,
    max_distance  = max_distance,
    pct_coastal   = pct_coastal,
    deepest_names = deepest_names,
    dist_counts   = dist_counts,
    dist_palette  = dist_palette,
    pal_fn        = pal_fn,
    cap_map       = cap_map,
    cap_dot       = cap_dot,
    crs_projected = spec$crs,
    label_long    = spec$label_long,
    unit_label    = spec$unit_label,
    gadm_link     = gadm_link,
    gisco_link    = gisco_link,
    igraph_link   = igraph_link,
    sf_link       = sf_link
  )
}

# ---- Combined Great Britain (England + Scotland + Wales) ---------------------
{
  gb_crs        <- 27700L
  gb_label_long <- "Great Britain's counties and council areas"
  gb_unit_label <- "county"

  gb_counties_raw <- rbind(
    regions[["England"]]$counties,
    regions[["Scotland"]]$counties,
    regions[["Wales"]]$counties
  )

  gb_counties <- .compute_sea_distance(gb_counties_raw, coastline,
                                       crs_projected = gb_crs)
  stopifnot(!any(is.na(gb_counties$sea_distance)))
  stopifnot(!any(is.na(gb_counties$county_name)))

  gb_summary    <- .summarise_sea_distance(gb_counties)
  gb_nb_table   <- .compute_neighbour_table(gb_counties, crs_projected = gb_crs)

  gb_n_counties    <- nrow(gb_counties)
  gb_n_coastal     <- sum(gb_counties$sea_distance == 0L, na.rm = TRUE)
  gb_n_inland      <- gb_n_counties - gb_n_coastal
  gb_max_distance  <- max(gb_counties$sea_distance, na.rm = TRUE)
  gb_pct_coastal   <- round(100 * gb_n_coastal / gb_n_counties, 1)
  gb_deepest_names <- sort(gb_counties$county_name[gb_counties$sea_distance == gb_max_distance])

  gb_dist_counts <- setNames(
    vapply(0:gb_max_distance,
           function(d) sum(gb_counties$sea_distance == d, na.rm = TRUE),
           integer(1L)),
    paste0("d", 0:gb_max_distance)
  )

  gb_dist_palette <- .build_palette(gb_max_distance)

  gb_pal_fn <- leaflet::colorFactor(
    palette = unname(gb_dist_palette[as.character(0:gb_max_distance)]),
    domain  = 0:gb_max_distance
  )

  gb_counties_wgs84 <- sf::st_transform(gb_counties, crs = 4326L)

  gb_cap_map <- paste0(
    "Interactive choropleth of ", gb_label_long, " coloured by sea-distance. ",
    "Includes cross-border neighbours: English counties touching Welsh or Scottish areas ",
    "are connected in the adjacency graph. ",
    "Sea-distance is the minimum number of shared-boundary hops to reach the coast ",
    "(distance 0 = directly coastal, 1 = one hop inland, etc.). ",
    "Of the ", gb_n_counties, " ", gb_unit_label, "s, ",
    gb_n_coastal, " (", gb_pct_coastal, "%) touch the coast directly ",
    "and ", gb_n_inland, " are inland. ",
    "The most inland (distance ", gb_max_distance, ") is ",
    paste(gb_deepest_names, collapse = " and "), ". ",
    "Click any area for details. ",
    "Source: ", gadm_link, " + ", gisco_link, " coastline."
  )

  gb_cap_dot <- paste0(
    "Dot plot showing the count of ", gb_label_long, " at each sea-distance level. ",
    "The x-axis gives the number of ", gb_unit_label, "s; ",
    "the y-axis gives sea distance in hops (0 = directly coastal). ",
    gb_n_coastal, " are coastal (distance 0); ",
    "the most inland (distance ", gb_max_distance, ") is ",
    paste(gb_deepest_names, collapse = " and "), ". ",
    "Source: ", gadm_link, "; sea distance via ", igraph_link, " BFS."
  )

  regions[["GreatBritain"]] <- list(
    counties       = gb_counties,
    counties_wgs84 = gb_counties_wgs84,
    summary        = gb_summary,
    nb_table       = gb_nb_table,
    n_counties     = gb_n_counties,
    n_coastal      = gb_n_coastal,
    n_inland       = gb_n_inland,
    max_distance   = gb_max_distance,
    pct_coastal    = gb_pct_coastal,
    deepest_names  = gb_deepest_names,
    dist_counts    = gb_dist_counts,
    dist_palette   = gb_dist_palette,
    pal_fn         = gb_pal_fn,
    cap_map        = gb_cap_map,
    cap_dot        = gb_cap_dot,
    crs_projected  = gb_crs,
    label_long     = gb_label_long,
    unit_label     = gb_unit_label,
    gadm_link      = gadm_link,
    gisco_link     = gisco_link,
    igraph_link    = igraph_link,
    sf_link        = sf_link
  )
}
```

```{=html}
<script>
document.addEventListener('DOMContentLoaded', function() {
  document.querySelectorAll('[data-bs-toggle="tab"]').forEach(function(tab) {
    tab.addEventListener('shown.bs.tab', function() {
      window.dispatchEvent(new Event('resize'));
    });
  });
  setTimeout(function() { window.dispatchEvent(new Event('resize')); }, 500);
  setTimeout(function() { window.dispatchEvent(new Event('resize')); }, 2000);
});
</script>
```

# Ireland

## Row

### {width="65%"}

```{r}
#| label: fig-map-ireland
#| fig-cap: !expr regions[["Ireland"]]$cap_map

.render_map(regions[["Ireland"]]$counties_wgs84, regions[["Ireland"]]$pal_fn)
```

### {width="35%"}

#### Key figures

```{r}
#| label: tbl-key-stats-ireland

.render_key_stats(regions[["Ireland"]])
```

#### About

Ireland has `r regions[["Ireland"]]$n_counties` traditional counties
(`r regions[["Ireland"]]$n_counties - 6L` in the Republic, 6 in Northern
Ireland). Of these, `r regions[["Ireland"]]$n_coastal`
(`r regions[["Ireland"]]$pct_coastal`%) touch the coast directly and
`r regions[["Ireland"]]$n_inland` are inland.

The most inland county (distance `r regions[["Ireland"]]$max_distance`) is
**`r paste(regions[["Ireland"]]$deepest_names, collapse = ", ")`**.

---

*Data: [GADM 4.1](https://gadm.org/) + [GISCO 2016](https://ec.europa.eu/eurostat/web/gisco).
Adjacency via [sf](https://r-spatial.github.io/sf/) with 200 m buffer in EPSG:2157;
BFS via [igraph](https://r.igraph.org/).*

## Row

### {width="40%"}

#### Sea-Distance Distribution

```{r}
#| label: tbl-summary-ireland

.render_summary_table(regions[["Ireland"]])
```

#### Sea-Distance Dot Plot

```{r}
#| label: fig-dot-ireland
#| fig-cap: !expr regions[["Ireland"]]$cap_dot
#| fig-height: 4.5

.render_dot_plot(regions[["Ireland"]])
```

### County Neighbours {width="60%" .full-width}

```{r}
#| label: tbl-neighbours-ireland

.render_neighbour_table(regions[["Ireland"]])
```

# England

## Row

### {width="65%"}

```{r}
#| label: fig-map-england
#| fig-cap: !expr regions[["England"]]$cap_map

.render_map(regions[["England"]]$counties_wgs84, regions[["England"]]$pal_fn)
```

### {width="35%"}

#### Key figures

```{r}
#| label: tbl-key-stats-england

.render_key_stats(regions[["England"]])
```

#### About

England has `r regions[["England"]]$n_counties` counties. Of these,
`r regions[["England"]]$n_coastal` (`r regions[["England"]]$pct_coastal`%)
touch the coast directly and `r regions[["England"]]$n_inland` are inland.

The most inland county (distance `r regions[["England"]]$max_distance`) is
**`r paste(regions[["England"]]$deepest_names, collapse = ", ")`**.

---

*Data: [GADM 4.1](https://gadm.org/) + [GISCO 2016](https://ec.europa.eu/eurostat/web/gisco).
Adjacency via [sf](https://r-spatial.github.io/sf/) with 200 m buffer in EPSG:27700;
BFS via [igraph](https://r.igraph.org/).*

## Row

### {width="40%"}

#### Sea-Distance Distribution

```{r}
#| label: tbl-summary-england

.render_summary_table(regions[["England"]])
```

#### Sea-Distance Dot Plot

```{r}
#| label: fig-dot-england
#| fig-cap: !expr regions[["England"]]$cap_dot
#| fig-height: 4.5

.render_dot_plot(regions[["England"]])
```

### County Neighbours {width="60%" .full-width}

```{r}
#| label: tbl-neighbours-england

.render_neighbour_table(regions[["England"]])
```

# Scotland

## Row

### {width="65%"}

```{r}
#| label: fig-map-scotland
#| fig-cap: !expr regions[["Scotland"]]$cap_map

.render_map(regions[["Scotland"]]$counties_wgs84, regions[["Scotland"]]$pal_fn)
```

### {width="35%"}

#### Key figures

```{r}
#| label: tbl-key-stats-scotland

.render_key_stats(regions[["Scotland"]])
```

#### About

Scotland has `r regions[["Scotland"]]$n_counties` council areas. Of these,
`r regions[["Scotland"]]$n_coastal` (`r regions[["Scotland"]]$pct_coastal`%)
touch the coast directly and `r regions[["Scotland"]]$n_inland` are inland.

The most inland council area (distance `r regions[["Scotland"]]$max_distance`) is
**`r paste(regions[["Scotland"]]$deepest_names, collapse = ", ")`**.

---

*Data: [GADM 4.1](https://gadm.org/) + [GISCO 2016](https://ec.europa.eu/eurostat/web/gisco).
Adjacency via [sf](https://r-spatial.github.io/sf/) with 200 m buffer in EPSG:27700;
BFS via [igraph](https://r.igraph.org/).*

## Row

### {width="40%"}

#### Sea-Distance Distribution

```{r}
#| label: tbl-summary-scotland

.render_summary_table(regions[["Scotland"]])
```

#### Sea-Distance Dot Plot

```{r}
#| label: fig-dot-scotland
#| fig-cap: !expr regions[["Scotland"]]$cap_dot
#| fig-height: 4.5

.render_dot_plot(regions[["Scotland"]])
```

### Council Area Neighbours {width="60%" .full-width}

```{r}
#| label: tbl-neighbours-scotland

.render_neighbour_table(regions[["Scotland"]])
```

# Wales

## Row

### {width="65%"}

```{r}
#| label: fig-map-wales
#| fig-cap: !expr regions[["Wales"]]$cap_map

.render_map(regions[["Wales"]]$counties_wgs84, regions[["Wales"]]$pal_fn)
```

### {width="35%"}

#### Key figures

```{r}
#| label: tbl-key-stats-wales

.render_key_stats(regions[["Wales"]])
```

#### About

Wales has `r regions[["Wales"]]$n_counties` counties. Of these,
`r regions[["Wales"]]$n_coastal` (`r regions[["Wales"]]$pct_coastal`%)
touch the coast directly and `r regions[["Wales"]]$n_inland` are inland.

The most inland county (distance `r regions[["Wales"]]$max_distance`) is
**`r paste(regions[["Wales"]]$deepest_names, collapse = ", ")`**.

---

*Data: [GADM 4.1](https://gadm.org/) + [GISCO 2016](https://ec.europa.eu/eurostat/web/gisco).
Adjacency via [sf](https://r-spatial.github.io/sf/) with 200 m buffer in EPSG:27700;
BFS via [igraph](https://r.igraph.org/).*

## Row

### {width="40%"}

#### Sea-Distance Distribution

```{r}
#| label: tbl-summary-wales

.render_summary_table(regions[["Wales"]])
```

#### Sea-Distance Dot Plot

```{r}
#| label: fig-dot-wales
#| fig-cap: !expr regions[["Wales"]]$cap_dot
#| fig-height: 4.5

.render_dot_plot(regions[["Wales"]])
```

### County Neighbours {width="60%" .full-width}

```{r}
#| label: tbl-neighbours-wales

.render_neighbour_table(regions[["Wales"]])
```

# Great Britain

## Row

### {width="65%"}

```{r}
#| label: fig-map-gb
#| fig-cap: !expr regions[["GreatBritain"]]$cap_map

.render_map(regions[["GreatBritain"]]$counties_wgs84, regions[["GreatBritain"]]$pal_fn)
```

### {width="35%"}

#### Key figures

```{r}
#| label: tbl-key-stats-gb

.render_key_stats(regions[["GreatBritain"]])
```

#### About

Great Britain combines England, Scotland, and Wales into a single adjacency
graph with `r regions[["GreatBritain"]]$n_counties` counties and council areas.
Cross-border neighbours — English counties touching Welsh or Scottish areas —
are included, providing information not available in the individual tabs.
Of these, `r regions[["GreatBritain"]]$n_coastal`
(`r regions[["GreatBritain"]]$pct_coastal`%) touch the coast directly and
`r regions[["GreatBritain"]]$n_inland` are inland.

The most inland area (distance `r regions[["GreatBritain"]]$max_distance`) is
**`r paste(regions[["GreatBritain"]]$deepest_names, collapse = ", ")`**.

---

*Data: [GADM 4.1](https://gadm.org/) + [GISCO 2016](https://ec.europa.eu/eurostat/web/gisco).
Adjacency via [sf](https://r-spatial.github.io/sf/) with 200 m buffer in EPSG:27700;
BFS via [igraph](https://r.igraph.org/).*

## Row

### {width="40%"}

#### Sea-Distance Distribution

```{r}
#| label: tbl-summary-gb

.render_summary_table(regions[["GreatBritain"]])
```

#### Sea-Distance Dot Plot

```{r}
#| label: fig-dot-gb
#| fig-cap: !expr regions[["GreatBritain"]]$cap_dot
#| fig-height: 4.5

.render_dot_plot(regions[["GreatBritain"]])
```

### County Neighbours {width="60%" .full-width}

```{r}
#| label: tbl-neighbours-gb

.render_neighbour_table(regions[["GreatBritain"]])
```
