---
title: "County Sea-Distance"
format:
  dashboard:
    orientation: rows
    scrolling: true
    theme:
      - cosmo
      - custom.scss
execute:
  echo: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false

library(sf)
library(leaflet)
library(dplyr)
library(igraph)
library(DT)

# ---- helpers (inline so the vignette is self-contained) ----------------------

.get_ireland_32_counties <- function() {
  irl <- sf::st_read(
    "https://geodata.ucdavis.edu/gadm/gadm4.1/json/gadm41_IRL_1.json",
    quiet = TRUE
  )
  irl$county_name <- ifelse(
    irl$HASC_1 == "IE.CK", "Cork", as.character(irl$NAME_1)
  )
  irl$country <- "Republic of Ireland"
  irl <- irl[, c("county_name", "country", "geometry")]

  gbr3 <- sf::st_read(
    "https://geodata.ucdavis.edu/gadm/gadm4.1/json/gadm41_GBR_3.json",
    quiet = TRUE
  )
  ni3 <- gbr3[!is.na(gbr3$NAME_1) & gbr3$NAME_1 == "NorthernIreland", ]

  ni_lookup <- c(
    Antrim = "Antrim", Newtownabbey = "Antrim", Ballymena = "Antrim",
    Ballymoney = "Antrim", Carrickfergus = "Antrim", Larne = "Antrim",
    Moyle = "Antrim", Belfast = "Antrim",
    Armagh = "Armagh", Banbridge = "Armagh", Craigavon = "Armagh",
    Ards = "Down", NorthDown = "Down", Castlereagh = "Down",
    Down = "Down", Lisburn = "Down", NewryandMourne = "Down",
    Fermanagh = "Fermanagh",
    Coleraine = "Londonderry", Derry = "Londonderry",
    Limavady = "Londonderry", Magherafelt = "Londonderry",
    Cookstown = "Tyrone", Dungannon = "Tyrone", Omagh = "Tyrone",
    Strabane = "Tyrone"
  )

  ni3$county_name <- ni_lookup[ni3$NAME_3]
  ni3 <- ni3[!is.na(ni3$county_name), ]

  ni_counties <- stats::aggregate(
    ni3["geometry"],
    by = list(county_name = ni3$county_name),
    FUN = function(x) sf::st_union(x)
  )
  ni_counties$country <- "Northern Ireland"
  ni_counties <- ni_counties[, c("county_name", "country", "geometry")]

  rbind(irl, ni_counties)
}

.compute_sea_distance <- function(counties, coastline, crs_projected = 27700L) {
  old_s2 <- sf::sf_use_s2()
  on.exit(sf::sf_use_s2(old_s2), add = TRUE)
  sf::sf_use_s2(FALSE)

  has_crs <- !is.na(sf::st_crs(counties))
  if (has_crs) {
    counties_m   <- sf::st_transform(counties, crs_projected)
    counties_buf <- sf::st_buffer(counties_m, 200)
    nb_mat       <- sf::st_intersects(counties_buf, counties_m, sparse = FALSE)
    diag(nb_mat) <- FALSE
    adj_matrix   <- nb_mat * 1L
  } else {
    adj_matrix <- sf::st_touches(counties, sparse = FALSE) * 1L
  }

  g <- igraph::graph_from_adjacency_matrix(adj_matrix, mode = "undirected",
                                           diag = FALSE)

  geom_types <- unique(as.character(sf::st_geometry_type(sf::st_geometry(coastline))))
  if (all(geom_types %in% c("POLYGON", "MULTIPOLYGON"))) {
    bbox_buf      <- sf::st_bbox(counties) + c(-1, -1, 1, 1)
    coast_cropped <- sf::st_crop(coastline, bbox_buf)
    coast_lines   <- sf::st_boundary(sf::st_geometry(coast_cropped))
    coastline_use <- sf::st_union(coast_lines)
  } else {
    coastline_use <- sf::st_union(sf::st_geometry(coastline))
  }

  if (has_crs && !is.na(sf::st_crs(coastline_use))) {
    coastline_use <- sf::st_transform(coastline_use, sf::st_crs(counties))
  }

  coastal     <- lengths(sf::st_intersects(counties, coastline_use)) > 0
  coastal_idx <- which(coastal)

  n         <- nrow(counties)
  min_dist  <- rep(Inf, n)

  if (length(coastal_idx) > 0 && length(coastal_idx) < n) {
    sp       <- igraph::distances(g, v = coastal_idx)
    min_dist <- apply(sp, 2, min)
  } else if (length(coastal_idx) == n) {
    min_dist <- rep(0, n)
  }

  # Island fix: counties that are coastal but appear disconnected get Inf
  # from the BFS (island counties have no land neighbours).
  # Reset any Inf that belongs to a directly-coastal county to 0.
  min_dist[is.infinite(min_dist) & coastal] <- 0
  dist <- as.integer(min_dist)

  counties$sea_distance <- dist
  counties
}

.compute_neighbour_table <- function(counties, crs_projected = 27700L) {
  old_s2 <- sf::sf_use_s2()
  on.exit(sf::sf_use_s2(old_s2), add = TRUE)
  sf::sf_use_s2(FALSE)

  counties_m <- sf::st_transform(counties, crs_projected)
  n <- nrow(counties_m)

  counties_buf <- sf::st_buffer(counties_m, 200)
  nb_mat <- sf::st_intersects(counties_buf, counties_m, sparse = FALSE)
  diag(nb_mat) <- FALSE

  results <- lapply(seq_len(n), function(i) {
    nb_idx   <- which(nb_mat[i, ])
    nb_names <- counties_m$county_name[nb_idx]

    cb_i     <- sf::st_boundary(sf::st_geometry(counties_m)[i])
    cb_i_buf <- sf::st_buffer(cb_i, 100)

    total_km <- sum(vapply(nb_idx, function(j) {
      cb_j   <- sf::st_boundary(sf::st_geometry(counties_m)[j])
      inters <- tryCatch(sf::st_intersection(cb_i_buf, cb_j), error = function(e) NULL)
      if (is.null(inters) || all(sf::st_is_empty(inters))) return(0)
      sum(as.numeric(sf::st_length(inters)), na.rm = TRUE) / 1000
    }, numeric(1L)))

    data.frame(
      county_name        = counties_m$county_name[i],
      n_neighbours       = length(nb_idx),
      neighbour_names    = paste(sort(nb_names), collapse = ", "),
      boundary_length_km = round(total_km, 1)
    )
  })

  out <- do.call(rbind, results)
  out[order(-out$n_neighbours, -out$boundary_length_km), ]
}

.summarise_sea_distance <- function(counties) {
  dplyr::count(sf::st_drop_geometry(counties), .data$sea_distance, name = "n")
}

# ---- render helpers ----------------------------------------------------------

.build_palette <- function(max_dist) {
  pal <- c("0" = "#fee090", "1" = "#fc8d59", "2" = "#d73027")
  if (max_dist > 2L) {
    extra <- scales::viridis_pal(option = "C")(max_dist - 2L)
    for (d in seq(3L, max_dist)) pal[as.character(d)] <- extra[d - 2L]
  }
  pal
}

.render_map <- function(counties_wgs84, pal_fn) {
  leaflet(data = counties_wgs84, height = "400px") |>
    leaflet::addProviderTiles(leaflet::providers$CartoDB.DarkMatter) |>
    leaflet::addPolygons(
      fillColor   = ~ pal_fn(sea_distance),
      fillOpacity = 0.8,
      color       = "#333333",
      weight      = 1.5,
      opacity     = 1,
      popup       = ~ paste0(
        "<strong>", county_name, "</strong><br>",
        country, "<br>",
        "Sea distance: ", sea_distance
      ),
      label       = ~ paste0(county_name, " (", sea_distance, ")")
    ) |>
    leaflet::addLegend(
      position  = "bottomleft",
      pal       = pal_fn,
      values    = ~ sea_distance,
      title     = "Sea distance",
      opacity   = 1,
      labFormat = leaflet::labelFormat(prefix = "Distance ")
    )
}

.render_key_stats <- function(r) {
  stats_df <- data.frame(
    Metric = c(
      "Total counties",
      "Coastal (distance 0)",
      paste0("Distance ", seq_len(r$max_distance),
             " (", seq_len(r$max_distance), " hop",
             ifelse(seq_len(r$max_distance) == 1L, "", "s"), ")"),
      "Maximum sea distance",
      "% directly coastal"
    ),
    Value = c(
      as.character(r$n_counties),
      as.character(r$dist_counts["d0"]),
      vapply(seq_len(r$max_distance),
             function(d) as.character(r$dist_counts[paste0("d", d)]),
             character(1L)),
      as.character(r$max_distance),
      paste0(r$pct_coastal, "%")
    )
  )
  knitr::kable(
    stats_df,
    align     = c("l", "r"),
    col.names = c("Metric", "Value"),
    caption   = paste0(
      "Summary statistics for ", r$label_long, ". ",
      r$n_coastal, " coastal, ", r$n_inland, " inland; ",
      paste(r$deepest_names, collapse = " and "),
      " is the most inland (distance ", r$max_distance, "). ",
      "Source: ", r$gadm_link, " + ", r$gisco_link, "."
    )
  )
}

.render_summary_table <- function(r) {
  df <- r$summary |>
    dplyr::mutate(
      pct   = round(100 * n / sum(n), 1),
      label = dplyr::case_when(
        sea_distance == 0L ~ "Coastal (0 hops)",
        sea_distance == 1L ~ "1 hop inland",
        TRUE               ~ paste0(sea_distance, " hops inland")
      )
    ) |>
    dplyr::select(label, n, pct)
  DT::datatable(
    df,
    colnames  = c("Sea distance", "Count", "Share (%)"),
    rownames  = FALSE,
    options   = list(dom = "t", paging = FALSE, ordering = TRUE),
    caption   = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; color: #ffffff;",
      paste0(
        "Distribution of ", r$label_long, ". ",
        r$n_coastal, " coastal, ", r$n_inland, " inland. ",
        paste(r$deepest_names, collapse = " and "),
        " is the most inland (distance ", r$max_distance, "). ",
        "Source: ", r$gadm_link, " + ", r$gisco_link,
        "; sea distance via ", r$igraph_link, " BFS."
      )
    )
  )
}

.render_dot_plot <- function(r) {
  par(mar = c(4, 8, 2, 1), bg = "black", fg = "white",
      col.axis = "white", col.lab = "white", col.main = "white")
  dotchart(
    x      = rev(r$summary$n),
    labels = rev(paste0("Distance ", r$summary$sea_distance)),
    pch    = 19,
    pt.cex = 2.5,
    color  = rev(r$dist_palette[as.character(r$summary$sea_distance)]),
    xlab   = paste0("Number of ", r$unit_label, "s"),
    main   = paste0("Sea-Distance Distribution (N = ", r$n_counties, ")"),
    lcolor = "grey40"
  )
  grid(col = "grey30", lty = 1)
}

.render_neighbour_table <- function(r) {
  df <- r$nb_table |>
    dplyr::select(county_name, n_neighbours, boundary_length_km, neighbour_names)
  DT::datatable(
    df,
    colnames  = c("County", "Neighbours", "Boundary (km)", "Neighbour names"),
    rownames  = FALSE,
    filter    = "top",
    options   = list(
      pageLength = 15,
      scrollX    = TRUE,
      order      = list(list(1, "desc")),
      columnDefs = list(
        list(width = "100px", targets = 0),
        list(width = "60px", targets = 1),
        list(width = "80px", targets = 2)
      )
    ),
    caption   = htmltools::tags$caption(
      style = "caption-side: top; text-align: left; color: #ffffff;",
      paste0(
        "All ", r$n_counties, " ", r$label_long,
        " sorted by neighbours (click headers to re-sort). ",
        "Use filters to search. ",
        "Boundary lengths in EPSG:", r$crs_projected, ". ",
        "Source: ", r$gadm_link, "; adjacency via ", r$sf_link, "."
      )
    )
  )
}

# ---- source links for captions -----------------------------------------------
gadm_link   <- "[GADM 4.1](https://gadm.org/data.html)"
gisco_link  <- "[GISCO 2016](https://ec.europa.eu/eurostat/web/gisco/geodata/statistical-units)"
igraph_link <- "[igraph](https://r.igraph.org/)"
sf_link     <- "[sf](https://r-spatial.github.io/sf/)"

# ---- fetch base data ---------------------------------------------------------
coastline <- giscoR::gisco_get_coastallines(year = 2016, resolution = "20")

ireland_counties_raw <- .get_ireland_32_counties()

gbr2 <- sf::st_read(
  "https://geodata.ucdavis.edu/gadm/gadm4.1/json/gadm41_GBR_2.json",
  quiet = TRUE
)
england_raw  <- gbr2[gbr2$NAME_1 == "England",  ]
scotland_raw <- gbr2[gbr2$NAME_1 == "Scotland", ]
wales_raw    <- gbr2[gbr2$NAME_1 == "Wales",    ]

for (obj_name in c("england_raw", "scotland_raw", "wales_raw")) {
  obj <- get(obj_name)
  obj$county_name <- as.character(obj$NAME_2)
  obj$country     <- as.character(obj$NAME_1)
  assign(obj_name, obj[, c("county_name", "country", "geometry")])
}

# ---- precompute all regions --------------------------------------------------
regions <- list()

region_specs <- list(
  Ireland  = list(raw = ireland_counties_raw,  crs = 2157L,  label_long = "Ireland's 32 traditional counties",  unit_label = "county"),
  England  = list(raw = england_raw,           crs = 27700L, label_long = "England's counties",                 unit_label = "county"),
  Scotland = list(raw = scotland_raw,          crs = 27700L, label_long = "Scotland's council areas",           unit_label = "council area"),
  Wales    = list(raw = wales_raw,             crs = 27700L, label_long = "Wales's counties",                   unit_label = "county")
)

for (rname in names(region_specs)) {
  spec      <- region_specs[[rname]]
  counties  <- .compute_sea_distance(spec$raw, coastline, crs_projected = spec$crs)
  stopifnot(!any(is.na(counties$sea_distance)))
  stopifnot(!any(is.na(counties$county_name)))

  summary_df  <- .summarise_sea_distance(counties)
  nb_table    <- .compute_neighbour_table(counties, crs_projected = spec$crs)

  n_counties   <- nrow(counties)
  n_coastal    <- sum(counties$sea_distance == 0L, na.rm = TRUE)
  n_inland     <- n_counties - n_coastal
  max_distance <- max(counties$sea_distance, na.rm = TRUE)
  pct_coastal  <- round(100 * n_coastal / n_counties, 1)
  deepest_names <- sort(counties$county_name[counties$sea_distance == max_distance])

  dist_counts <- setNames(
    vapply(0:max_distance,
           function(d) sum(counties$sea_distance == d, na.rm = TRUE),
           integer(1L)),
    paste0("d", 0:max_distance)
  )

  dist_palette <- .build_palette(max_distance)

  pal_fn <- leaflet::colorFactor(
    palette = unname(dist_palette[as.character(0:max_distance)]),
    domain  = 0:max_distance
  )

  counties_wgs84 <- sf::st_transform(counties, crs = 4326L)

  cap_map <- paste0(
    "Interactive choropleth of ", spec$label_long, " coloured by sea-distance. ",
    "Sea-distance is the minimum number of shared-boundary hops to reach the coast ",
    "(distance 0 = directly coastal, 1 = one hop inland, etc.). ",
    "Of the ", n_counties, " ", spec$unit_label, "s, ",
    n_coastal, " (", pct_coastal, "%) touch the coast directly ",
    "and ", n_inland, " are inland. ",
    "The most inland (distance ", max_distance, ") is ",
    paste(deepest_names, collapse = " and "), ". ",
    "Click any area for details. ",
    "Source: ", gadm_link, " + ", gisco_link, " coastline."
  )

  cap_dot <- paste0(
    "Dot plot showing the count of ", spec$label_long, " at each sea-distance level. ",
    "The x-axis gives the number of ", spec$unit_label, "s; ",
    "the y-axis gives sea distance in hops (0 = directly coastal). ",
    n_coastal, " are coastal (distance 0); ",
    "the most inland (distance ", max_distance, ") is ",
    paste(deepest_names, collapse = " and "), ". ",
    "Source: ", gadm_link, "; sea distance via ", igraph_link, " BFS."
  )

  regions[[rname]] <- list(
    counties      = counties,
    counties_wgs84 = counties_wgs84,
    summary       = summary_df,
    nb_table      = nb_table,
    n_counties    = n_counties,
    n_coastal     = n_coastal,
    n_inland      = n_inland,
    max_distance  = max_distance,
    pct_coastal   = pct_coastal,
    deepest_names = deepest_names,
    dist_counts   = dist_counts,
    dist_palette  = dist_palette,
    pal_fn        = pal_fn,
    cap_map       = cap_map,
    cap_dot       = cap_dot,
    crs_projected = spec$crs,
    label_long    = spec$label_long,
    unit_label    = spec$unit_label,
    gadm_link     = gadm_link,
    gisco_link    = gisco_link,
    igraph_link   = igraph_link,
    sf_link       = sf_link
  )
}
```

```{=html}
<script>
document.addEventListener('DOMContentLoaded', function() {
  document.querySelectorAll('[data-bs-toggle="tab"]').forEach(function(tab) {
    tab.addEventListener('shown.bs.tab', function() {
      window.dispatchEvent(new Event('resize'));
    });
  });
  setTimeout(function() { window.dispatchEvent(new Event('resize')); }, 500);
  setTimeout(function() { window.dispatchEvent(new Event('resize')); }, 2000);
});
</script>
```

# Ireland

## Row

### {width="65%"}

```{r}
#| label: fig-map-ireland
#| fig-cap: !expr regions[["Ireland"]]$cap_map

.render_map(regions[["Ireland"]]$counties_wgs84, regions[["Ireland"]]$pal_fn)
```

### {width="35%"}

#### Key figures

```{r}
#| label: tbl-key-stats-ireland

.render_key_stats(regions[["Ireland"]])
```

#### About

Ireland has `r regions[["Ireland"]]$n_counties` traditional counties
(`r regions[["Ireland"]]$n_counties - 6L` in the Republic, 6 in Northern
Ireland). Of these, `r regions[["Ireland"]]$n_coastal`
(`r regions[["Ireland"]]$pct_coastal`%) touch the coast directly and
`r regions[["Ireland"]]$n_inland` are inland.

The most inland county (distance `r regions[["Ireland"]]$max_distance`) is
**`r paste(regions[["Ireland"]]$deepest_names, collapse = ", ")`**.

---

*Data: [GADM 4.1](https://gadm.org/) + [GISCO 2016](https://ec.europa.eu/eurostat/web/gisco).
Adjacency via [sf](https://r-spatial.github.io/sf/) with 200 m buffer in EPSG:2157;
BFS via [igraph](https://r.igraph.org/).*

## Row

### {width="40%"}

#### Sea-Distance Distribution

```{r}
#| label: tbl-summary-ireland

.render_summary_table(regions[["Ireland"]])
```

#### Sea-Distance Dot Plot

```{r}
#| label: fig-dot-ireland
#| fig-cap: !expr regions[["Ireland"]]$cap_dot
#| fig-height: 1.5

.render_dot_plot(regions[["Ireland"]])
```

### County Neighbours {width="60%" .full-width}

```{r}
#| label: tbl-neighbours-ireland

.render_neighbour_table(regions[["Ireland"]])
```

# England

## Row

### {width="65%"}

```{r}
#| label: fig-map-england
#| fig-cap: !expr regions[["England"]]$cap_map

.render_map(regions[["England"]]$counties_wgs84, regions[["England"]]$pal_fn)
```

### {width="35%"}

#### Key figures

```{r}
#| label: tbl-key-stats-england

.render_key_stats(regions[["England"]])
```

#### About

England has `r regions[["England"]]$n_counties` counties. Of these,
`r regions[["England"]]$n_coastal` (`r regions[["England"]]$pct_coastal`%)
touch the coast directly and `r regions[["England"]]$n_inland` are inland.

The most inland county (distance `r regions[["England"]]$max_distance`) is
**`r paste(regions[["England"]]$deepest_names, collapse = ", ")`**.

---

*Data: [GADM 4.1](https://gadm.org/) + [GISCO 2016](https://ec.europa.eu/eurostat/web/gisco).
Adjacency via [sf](https://r-spatial.github.io/sf/) with 200 m buffer in EPSG:27700;
BFS via [igraph](https://r.igraph.org/).*

## Row

### {width="40%"}

#### Sea-Distance Distribution

```{r}
#| label: tbl-summary-england

.render_summary_table(regions[["England"]])
```

#### Sea-Distance Dot Plot

```{r}
#| label: fig-dot-england
#| fig-cap: !expr regions[["England"]]$cap_dot
#| fig-height: 1.5

.render_dot_plot(regions[["England"]])
```

### County Neighbours {width="60%" .full-width}

```{r}
#| label: tbl-neighbours-england

.render_neighbour_table(regions[["England"]])
```

# Scotland

## Row

### {width="65%"}

```{r}
#| label: fig-map-scotland
#| fig-cap: !expr regions[["Scotland"]]$cap_map

.render_map(regions[["Scotland"]]$counties_wgs84, regions[["Scotland"]]$pal_fn)
```

### {width="35%"}

#### Key figures

```{r}
#| label: tbl-key-stats-scotland

.render_key_stats(regions[["Scotland"]])
```

#### About

Scotland has `r regions[["Scotland"]]$n_counties` council areas. Of these,
`r regions[["Scotland"]]$n_coastal` (`r regions[["Scotland"]]$pct_coastal`%)
touch the coast directly and `r regions[["Scotland"]]$n_inland` are inland.

The most inland council area (distance `r regions[["Scotland"]]$max_distance`) is
**`r paste(regions[["Scotland"]]$deepest_names, collapse = ", ")`**.

---

*Data: [GADM 4.1](https://gadm.org/) + [GISCO 2016](https://ec.europa.eu/eurostat/web/gisco).
Adjacency via [sf](https://r-spatial.github.io/sf/) with 200 m buffer in EPSG:27700;
BFS via [igraph](https://r.igraph.org/).*

## Row

### {width="40%"}

#### Sea-Distance Distribution

```{r}
#| label: tbl-summary-scotland

.render_summary_table(regions[["Scotland"]])
```

#### Sea-Distance Dot Plot

```{r}
#| label: fig-dot-scotland
#| fig-cap: !expr regions[["Scotland"]]$cap_dot
#| fig-height: 1.5

.render_dot_plot(regions[["Scotland"]])
```

### Council Area Neighbours {width="60%" .full-width}

```{r}
#| label: tbl-neighbours-scotland

.render_neighbour_table(regions[["Scotland"]])
```

# Wales

## Row

### {width="65%"}

```{r}
#| label: fig-map-wales
#| fig-cap: !expr regions[["Wales"]]$cap_map

.render_map(regions[["Wales"]]$counties_wgs84, regions[["Wales"]]$pal_fn)
```

### {width="35%"}

#### Key figures

```{r}
#| label: tbl-key-stats-wales

.render_key_stats(regions[["Wales"]])
```

#### About

Wales has `r regions[["Wales"]]$n_counties` counties. Of these,
`r regions[["Wales"]]$n_coastal` (`r regions[["Wales"]]$pct_coastal`%)
touch the coast directly and `r regions[["Wales"]]$n_inland` are inland.

The most inland county (distance `r regions[["Wales"]]$max_distance`) is
**`r paste(regions[["Wales"]]$deepest_names, collapse = ", ")`**.

---

*Data: [GADM 4.1](https://gadm.org/) + [GISCO 2016](https://ec.europa.eu/eurostat/web/gisco).
Adjacency via [sf](https://r-spatial.github.io/sf/) with 200 m buffer in EPSG:27700;
BFS via [igraph](https://r.igraph.org/).*

## Row

### {width="40%"}

#### Sea-Distance Distribution

```{r}
#| label: tbl-summary-wales

.render_summary_table(regions[["Wales"]])
```

#### Sea-Distance Dot Plot

```{r}
#| label: fig-dot-wales
#| fig-cap: !expr regions[["Wales"]]$cap_dot
#| fig-height: 1.5

.render_dot_plot(regions[["Wales"]])
```

### County Neighbours {width="60%" .full-width}

```{r}
#| label: tbl-neighbours-wales

.render_neighbour_table(regions[["Wales"]])
```
