---
title: "Ireland County Sea-Distance Dashboard"
format:
  dashboard:
    orientation: pages
    scrolling: false
    theme: cosmo
    css: ../../styles.css
execute:
  echo: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false

library(sf)
library(leaflet)
library(dplyr)
library(giscoR)
library(igraph)

# ---- helpers (inline so the vignette is self-contained) ----------------------

.get_ireland_boundaries <- function(year = 2021) {
  nuts <- giscoR::gisco_get_nuts(
    year = year,
    resolution = "10",
    nuts_level = 3L,
    country = c("IE", "UK")
  )
  nuts[grepl("^IE|^UKN", nuts$NUTS_ID), ]
}

.get_coastline <- function(year = 2021) {
  giscoR::gisco_get_coastallines(year = year, resolution = "10")
}

.compute_sea_distance <- function(counties, coastline) {
  adj_matrix <- sf::st_touches(counties, sparse = FALSE) * 1L
  g <- igraph::graph_from_adjacency_matrix(adj_matrix, mode = "undirected",
                                           diag = FALSE)
  coastal    <- lengths(sf::st_intersects(counties, coastline)) > 0
  coastal_idx <- which(coastal)
  n          <- nrow(counties)
  dist       <- rep(NA_integer_, n)

  if (length(coastal_idx) > 0 && length(coastal_idx) < n) {
    sp       <- igraph::distances(g, v = coastal_idx)
    dist     <- as.integer(apply(sp, 2, min))
  } else if (length(coastal_idx) == n) {
    dist     <- rep(0L, n)
  }
  counties$sea_distance <- dist
  counties
}

.summarise_sea_distance <- function(counties) {
  dplyr::count(sf::st_drop_geometry(counties), .data$sea_distance, name = "n")
}

# ---- fetch data ---------------------------------------------------------------
counties  <- .get_ireland_boundaries()
coastline <- .get_coastline()
counties  <- .compute_sea_distance(counties, coastline)
summary_df <- .summarise_sea_distance(counties)

# ---- derived values for dynamic prose ----------------------------------------
n_counties    <- nrow(counties)
n_coastal     <- sum(counties$sea_distance == 0L, na.rm = TRUE)
n_one_hop     <- sum(counties$sea_distance == 1L, na.rm = TRUE)
n_two_hop     <- sum(counties$sea_distance == 2L, na.rm = TRUE)
max_distance  <- max(counties$sea_distance, na.rm = TRUE)
pct_coastal   <- round(100 * n_coastal / n_counties, 1)

# ---- colour palette (viridis, colorblind-safe) --------------------------------
dist_levels   <- sort(unique(counties$sea_distance))
pal_viridis   <- c("0" = "#440154", "1" = "#31688e", "2" = "#35b779")
# Extend automatically if max_distance > 2
all_levels    <- as.character(seq_len(max_distance + 1L) - 1L)
viridis_ramp  <- scales::viridis_pal(option = "D")(length(all_levels))
names(viridis_ramp) <- all_levels

pal_fn <- leaflet::colorFactor(
  palette = viridis_ramp,
  domain  = as.character(counties$sea_distance)
)

counties_wgs84 <- sf::st_transform(counties, crs = 4326)
```

# Ireland Map {orientation="columns"}

## Column {width="70%"}

```{r}
#| label: fig-map
#| fig-cap: !expr paste0(
#|   "Interactive choropleth of Ireland's ", n_counties, " NUTS3 regions coloured by sea-distance — ",
#|   "the minimum number of shared-boundary county-hops required to reach the coast ",
#|   "(distance 0 = directly coastal, 1 = one hop inland, 2 = two hops inland). ",
#|   "Colour scale follows the viridis palette (purple → teal → green) for ",
#|   "colorblind accessibility; distance categories are labelled in the map legend. ",
#|   "Of the ", n_counties, " regions shown, ", n_coastal, " (", pct_coastal, "%) ",
#|   "touch the coast directly, reflecting Ireland's highly indented Atlantic and Irish Sea coastline. ",
#|   "Click any region to see its NUTS3 name and sea-distance value. ",
#|   "Source: GISCO NUTS 2021 (IE + UKN regions) fetched via the giscoR package."
#| )

leaflet(data = counties_wgs84) |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addPolygons(
    fillColor   = ~ pal_fn(as.character(sea_distance)),
    fillOpacity = 0.75,
    color       = "#ffffff",
    weight      = 1,
    opacity     = 1,
    popup       = ~ paste0(
      "<strong>", NUTS_NAME, "</strong><br>",
      "NUTS ID: ", NUTS_ID, "<br>",
      "Sea distance: ", sea_distance
    ),
    label       = ~ paste0(NUTS_NAME, " (", sea_distance, ")")
  ) |>
  leaflet::addLegend(
    position = "bottomright",
    pal      = pal_fn,
    values   = ~ as.character(sea_distance),
    title    = "Sea distance<br>(county hops)",
    opacity  = 0.85
  )
```

## Column {width="30%"}

### Key figures

```{r}
#| label: tbl-key-stats

knitr::kable(
  data.frame(
    Metric = c(
      "Total NUTS3 regions",
      "Coastal (distance 0)",
      "One hop inland (distance 1)",
      "Two hops inland (distance 2)",
      "Maximum sea distance",
      "% directly coastal"
    ),
    Value = c(
      n_counties,
      n_coastal,
      n_one_hop,
      n_two_hop,
      max_distance,
      paste0(pct_coastal, "%")
    )
  ),
  align = c("l", "r"),
  col.names = c("Metric", "Value"),
  caption = paste0(
    "Summary statistics for Ireland's ", n_counties, " NUTS3 regions. ",
    "Sea distance is computed via breadth-first search on a county-adjacency graph, ",
    "where coastal counties (those intersecting the GISCO coastline geometry) are ",
    "assigned distance 0 and each subsequent hop adds 1. ",
    "The table highlights that ", pct_coastal, "% of regions touch the coast directly, ",
    "consistent with Ireland's elongated and deeply indented coastline. ",
    "Source: GISCO NUTS 2021 (giscoR), coastline from gisco_get_coastallines()."
  )
)
```

### About the sea-distance metric

Ireland has `r n_counties` NUTS3 regions. Sea distance measures how many
county-boundaries separate a region from the nearest coastal region.
Coastal regions (distance = 0) border the sea directly. Of these,
`r n_one_hop` region(s) are one county-hop from the coast, and
`r n_two_hop` region(s) are two hops inland.

The maximum sea distance across the island is **`r max_distance`**,
reflecting Ireland's compact geography relative to its coastline length.

---

*Data: GISCO NUTS 2021 via [giscoR](https://ropengov.github.io/giscoR/).
Adjacency computed with `sf::st_touches()`; BFS via
[igraph](https://r.igraph.org/).*

# Summary {orientation="columns"}

## Column {width="50%"}

```{r}
#| label: tbl-summary

knitr::kable(
  summary_df |>
    dplyr::mutate(
      pct = round(100 * n / sum(n), 1),
      label = dplyr::case_when(
        sea_distance == 0L ~ "Coastal (0 hops)",
        sea_distance == 1L ~ "One hop inland",
        sea_distance == 2L ~ "Two hops inland",
        TRUE               ~ paste0(sea_distance, " hops inland")
      )
    ) |>
    dplyr::select(label, n, pct) |>
    dplyr::rename(
      "Sea distance"       = label,
      "Count"              = n,
      "Share (%)"          = pct
    ),
  align   = c("l", "r", "r"),
  caption = paste0(
    "Distribution of Ireland's ", n_counties, " NUTS3 regions across sea-distance categories. ",
    "Each row corresponds to one distance level (0 = coastal, 1 = one shared-boundary hop, etc.). ",
    "'Share' is the row's count expressed as a percentage of all ", n_counties, " regions. ",
    "The table shows that coastal and near-coastal regions dominate: ",
    n_coastal + n_one_hop, " out of ", n_counties, " regions lie within one hop of the sea (",
    round(100 * (n_coastal + n_one_hop) / n_counties, 1), "%). ",
    "Source: GISCO NUTS 2021 fetched via giscoR; sea distance computed with igraph BFS."
  )
)
```

## Column {width="50%"}

```{r}
#| label: fig-bar
#| fig-cap: !expr paste0(
#|   "Bar chart showing the count of Ireland's ", n_counties, " NUTS3 regions at each sea-distance level. ",
#|   "The x-axis gives sea distance in county hops (0 = directly coastal); ",
#|   "the y-axis gives the number of NUTS3 regions at that distance. ",
#|   "Bars are coloured using the same viridis palette as the map on Page 1 to enable ",
#|   "direct visual correspondence between the two views. ",
#|   "The distribution is strongly right-skewed: ", n_coastal, " regions are coastal ",
#|   "and counts fall rapidly with increasing inland distance, ",
#|   "reflecting Ireland's elongated shape and the fact that no point on the island is ",
#|   "more than ", max_distance, " county hops from the sea. ",
#|   "Source: GISCO NUTS 2021 (giscoR); sea distance from igraph BFS."
#| )

barplot(
  height   = summary_df$n,
  names.arg = paste0(summary_df$sea_distance, "\nhop", ifelse(summary_df$sea_distance == 1L, "", "s")),
  col      = viridis_ramp[as.character(summary_df$sea_distance)],
  border   = NA,
  xlab     = "Sea distance (county hops)",
  ylab     = "Number of NUTS3 regions",
  main     = paste0("Sea-Distance Distribution (N = ", n_counties, " regions)"),
  las      = 1
)
```
