---
title: "County Sea-Distance"
format:
  dashboard:
    orientation: rows
    scrolling: true
    theme:
      - cosmo
      - custom.scss
execute:
  echo: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false

library(sf)
library(leaflet)
library(dplyr)
library(igraph)

# ---- helpers (inline so the vignette is self-contained) ----------------------

.get_gb_counties <- function() {
  gbr2 <- sf::st_read(
    "https://geodata.ucdavis.edu/gadm/gadm4.1/json/gadm41_GBR_2.json",
    quiet = TRUE
  )
  gb <- gbr2[gbr2$NAME_1 %in% c("England", "Scotland", "Wales"), ]
  gb$county_name <- as.character(gb$NAME_2)
  gb$country     <- as.character(gb$NAME_1)
  gb[, c("county_name", "country", "geometry")]
}

.compute_sea_distance <- function(counties, coastline, crs_projected = 27700L) {
  old_s2 <- sf::sf_use_s2()
  on.exit(sf::sf_use_s2(old_s2), add = TRUE)
  sf::sf_use_s2(FALSE)

  has_crs <- !is.na(sf::st_crs(counties))
  if (has_crs) {
    counties_m   <- sf::st_transform(counties, crs_projected)
    counties_buf <- sf::st_buffer(counties_m, 200)
    nb_mat       <- sf::st_intersects(counties_buf, counties_m, sparse = FALSE)
    diag(nb_mat) <- FALSE
    adj_matrix   <- nb_mat * 1L
  } else {
    adj_matrix <- sf::st_touches(counties, sparse = FALSE) * 1L
  }

  g <- igraph::graph_from_adjacency_matrix(adj_matrix, mode = "undirected",
                                           diag = FALSE)

  geom_types <- unique(as.character(sf::st_geometry_type(sf::st_geometry(coastline))))
  if (all(geom_types %in% c("POLYGON", "MULTIPOLYGON"))) {
    bbox_buf      <- sf::st_bbox(counties) + c(-1, -1, 1, 1)
    coast_cropped <- sf::st_crop(coastline, bbox_buf)
    coast_lines   <- sf::st_boundary(sf::st_geometry(coast_cropped))
    coastline_use <- sf::st_union(coast_lines)
  } else {
    coastline_use <- sf::st_union(sf::st_geometry(coastline))
  }

  if (has_crs && !is.na(sf::st_crs(coastline_use))) {
    coastline_use <- sf::st_transform(coastline_use, sf::st_crs(counties))
  }

  coastal     <- lengths(sf::st_intersects(counties, coastline_use)) > 0
  coastal_idx <- which(coastal)

  n    <- nrow(counties)
  dist <- rep(NA_integer_, n)
  dist[coastal_idx] <- 0L

  if (length(coastal_idx) > 0 && length(coastal_idx) < n) {
    sp   <- igraph::distances(g, v = coastal_idx)
    dist <- as.integer(apply(sp, 2, min))
  } else if (length(coastal_idx) == n) {
    dist <- rep(0L, n)
  }

  counties$sea_distance <- dist
  counties
}

.compute_neighbour_table <- function(counties, crs_projected = 27700L) {
  old_s2 <- sf::sf_use_s2()
  on.exit(sf::sf_use_s2(old_s2), add = TRUE)
  sf::sf_use_s2(FALSE)

  counties_m <- sf::st_transform(counties, crs_projected)
  n <- nrow(counties_m)

  counties_buf <- sf::st_buffer(counties_m, 200)
  nb_mat <- sf::st_intersects(counties_buf, counties_m, sparse = FALSE)
  diag(nb_mat) <- FALSE

  results <- lapply(seq_len(n), function(i) {
    nb_idx   <- which(nb_mat[i, ])
    nb_names <- counties_m$county_name[nb_idx]

    cb_i     <- sf::st_boundary(sf::st_geometry(counties_m)[i])
    cb_i_buf <- sf::st_buffer(cb_i, 100)

    total_km <- sum(vapply(nb_idx, function(j) {
      cb_j   <- sf::st_boundary(sf::st_geometry(counties_m)[j])
      inters <- tryCatch(sf::st_intersection(cb_i_buf, cb_j), error = function(e) NULL)
      if (is.null(inters) || all(sf::st_is_empty(inters))) return(0)
      sum(as.numeric(sf::st_length(inters)), na.rm = TRUE) / 1000
    }, numeric(1L)))

    data.frame(
      county_name        = counties_m$county_name[i],
      country            = counties_m$country[i],
      n_neighbours       = length(nb_idx),
      neighbour_names    = paste(sort(nb_names), collapse = ", "),
      boundary_length_km = round(total_km, 1)
    )
  })

  out <- do.call(rbind, results)
  out[order(-out$n_neighbours, -out$boundary_length_km), ]
}

.summarise_sea_distance <- function(counties) {
  dplyr::count(sf::st_drop_geometry(counties), .data$sea_distance, name = "n")
}

# ---- fetch data ---------------------------------------------------------------
counties   <- .get_gb_counties()
coastline  <- giscoR::gisco_get_coastallines(year = 2016, resolution = "03")
counties   <- .compute_sea_distance(counties, coastline, crs_projected = 27700L)
summary_df <- .summarise_sea_distance(counties)
nb_table   <- .compute_neighbour_table(counties, crs_projected = 27700L)

# ---- derived values for dynamic prose ----------------------------------------
n_counties    <- nrow(counties)
n_coastal     <- sum(counties$sea_distance == 0L, na.rm = TRUE)
n_inland      <- n_counties - n_coastal
max_distance  <- max(counties$sea_distance, na.rm = TRUE)
pct_coastal   <- round(100 * n_coastal / n_counties, 1)

coastal_names <- sort(counties$county_name[counties$sea_distance == 0L])
inland_names  <- sort(counties$county_name[counties$sea_distance > 0L])
deepest_names <- sort(counties$county_name[counties$sea_distance == max_distance])

dist_counts <- setNames(
  vapply(0:max_distance, function(d) sum(counties$sea_distance == d, na.rm = TRUE), integer(1)),
  paste0("d", 0:max_distance)
)

# ---- source links for captions -----------------------------------------------
gadm_link   <- "[GADM 4.1](https://gadm.org/data.html)"
gisco_link  <- "[GISCO 2016](https://ec.europa.eu/eurostat/web/gisco/geodata/statistical-units)"
igraph_link <- "[igraph](https://r.igraph.org/)"
sf_link     <- "[sf](https://r-spatial.github.io/sf/)"

# ---- pre-computed captions ---------------------------------------------------
cap_map <- paste0(
  "Interactive choropleth of Great Britain's ", n_counties, " counties coloured by sea-distance. ",
  "Sea-distance is the minimum number of shared-boundary county-hops to reach the coast ",
  "(distance 0 = directly coastal, 1 = one hop inland, etc.). ",
  "Of the ", n_counties, " counties, ", n_coastal, " (", pct_coastal, "%) touch the coast directly ",
  "and ", n_inland, " are inland. ",
  "The most inland county (distance ", max_distance, ") is ",
  paste(deepest_names, collapse = " and "), ". ",
  "Click any county for details. ",
  "Source: ", gadm_link, " + ", gisco_link, " coastline."
)

cap_dot <- paste0(
  "Dot plot showing the count of Great Britain's ", n_counties, " counties ",
  "at each sea-distance level. ",
  "The x-axis gives the number of counties; ",
  "the y-axis gives sea distance in county hops (0 = directly coastal). ",
  n_coastal, " counties are coastal (distance 0); ",
  "the most inland (distance ", max_distance, ") is ",
  paste(deepest_names, collapse = " and "), ". ",
  "Source: ", gadm_link, "; sea distance via ", igraph_link, " BFS."
)

# ---- colour palette (high-contrast on black DarkMatter tiles) ----------------
dist_palette <- c("0" = "#fee090", "1" = "#fc8d59", "2" = "#d73027")
if (max_distance > 2L) {
  extra <- scales::viridis_pal(option = "C")(max_distance - 2L)
  for (d in 3:max_distance) dist_palette[as.character(d)] <- extra[d - 2L]
}

pal_fn <- leaflet::colorFactor(
  palette = unname(dist_palette[as.character(0:max_distance)]),
  domain  = 0:max_distance
)

counties_wgs84 <- sf::st_transform(counties, crs = 4326)
```

# Great Britain {orientation="columns"}

## Column {width="65%"}

```{r}
#| label: fig-map
#| fig-cap: !expr cap_map

leaflet(data = counties_wgs84, height = "600px") |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.DarkMatter) |>
  leaflet::addPolygons(
    fillColor   = ~ pal_fn(sea_distance),
    fillOpacity = 0.8,
    color       = "#333333",
    weight      = 1.5,
    opacity     = 1,
    popup       = ~ paste0(
      "<strong>", county_name, "</strong><br>",
      country, "<br>",
      "Sea distance: ", sea_distance
    ),
    label       = ~ paste0(county_name, " (", sea_distance, ")")
  ) |>
  leaflet::addLegend(
    position = "bottomleft",
    pal      = pal_fn,
    values   = ~ sea_distance,
    title    = "Sea distance",
    opacity  = 1,
    labFormat = leaflet::labelFormat(prefix = "Distance ")
  )
```

```{=html}
<script>
document.addEventListener('DOMContentLoaded', function() {
  document.querySelectorAll('[data-bs-toggle="tab"]').forEach(function(tab) {
    tab.addEventListener('shown.bs.tab', function() {
      window.dispatchEvent(new Event('resize'));
    });
  });
  setTimeout(function() { window.dispatchEvent(new Event('resize')); }, 500);
  setTimeout(function() { window.dispatchEvent(new Event('resize')); }, 2000);
});
</script>
```

## Column {width="35%"}

### Key figures

```{r}
#| label: tbl-key-stats

stats_df <- data.frame(
  Metric = c(
    "Total counties",
    paste0("Coastal (distance 0)"),
    paste0("Distance ", seq_len(max_distance), " (", seq_len(max_distance), " hop",
           ifelse(seq_len(max_distance) == 1, "", "s"), ")"),
    "Maximum sea distance",
    "% directly coastal"
  ),
  Value = c(
    as.character(n_counties),
    as.character(dist_counts["d0"]),
    vapply(seq_len(max_distance), function(d) as.character(dist_counts[paste0("d", d)]), character(1)),
    as.character(max_distance),
    paste0(pct_coastal, "%")
  )
)
knitr::kable(stats_df, align = c("l", "r"), col.names = c("Metric", "Value"),
  caption = paste0(
    "Summary statistics for Great Britain's ", n_counties, " counties (England, Scotland, Wales). ",
    n_coastal, " coastal, ", n_inland, " inland; ",
    paste(deepest_names, collapse = " and "), " is the most inland (distance ", max_distance, "). ",
    "Projected in EPSG:27700 (British National Grid). ",
    "Source: ", gadm_link, " + ", gisco_link, "."
  )
)
```

### About

Great Britain has `r n_counties` counties across England, Scotland, and Wales
(Northern Ireland is excluded — see the Ireland dashboard).
Of these, `r n_coastal` (`r pct_coastal`%) touch the coast directly
and `r n_inland` are inland.

The most inland county (distance `r max_distance`) is
**`r paste(deepest_names, collapse = ", ")`**.

---

*Data: [`r gadm_link`](https://gadm.org/) + [`r gisco_link`](https://ec.europa.eu/eurostat/web/gisco).
Adjacency via [`r sf_link`](https://r-spatial.github.io/sf/) with 200 m buffer in EPSG:27700;
BFS via [`r igraph_link`](https://r.igraph.org/).*

# Details {orientation="rows"}

## Row {height="45%"}

### Sea-Distance Distribution

```{r}
#| label: tbl-summary

knitr::kable(
  summary_df |>
    dplyr::mutate(
      pct = round(100 * n / sum(n), 1),
      label = dplyr::case_when(
        sea_distance == 0L ~ "Coastal (0 hops)",
        sea_distance == 1L ~ "1 hop inland",
        TRUE               ~ paste0(sea_distance, " hops inland")
      )
    ) |>
    dplyr::select(label, n, pct) |>
    dplyr::rename("Sea distance" = label, "Count" = n, "Share (%)" = pct),
  align = c("l", "r", "r"),
  caption = paste0(
    "Distribution of Great Britain's ", n_counties, " counties across sea-distance categories. ",
    n_coastal, " are coastal, ", n_inland, " are inland. ",
    paste(deepest_names, collapse = " and "), " is the only distance-", max_distance, " county. ",
    "Source: ", gadm_link, " + ", gisco_link, "; sea distance via ", igraph_link, " BFS."
  )
)
```

### Sea-Distance Dot Plot

```{r}
#| label: fig-dot
#| fig-cap: !expr cap_dot
#| fig-height: 3

par(mar = c(4, 8, 2, 1), bg = "black", fg = "white",
    col.axis = "white", col.lab = "white", col.main = "white")
dotchart(
  x      = rev(summary_df$n),
  labels = rev(paste0("Distance ", summary_df$sea_distance)),
  pch    = 19,
  pt.cex = 2.5,
  color  = rev(dist_palette[as.character(summary_df$sea_distance)]),
  xlab   = "Number of counties",
  main   = paste0("Sea-Distance Distribution (N = ", n_counties, ")"),
  lcolor = "grey40"
)
grid(col = "grey30", lty = 1)
```

## Row {height="55%"}

### County Neighbours {.full-width}

```{r}
#| label: tbl-neighbours

knitr::kable(
  nb_table |>
    dplyr::select(county_name, country, n_neighbours, boundary_length_km, neighbour_names) |>
    dplyr::rename(
      "County"         = county_name,
      "Country"        = country,
      "Neighbours"     = n_neighbours,
      "Boundary (km)"  = boundary_length_km,
      "Neighbour names" = neighbour_names
    ),
  align = c("l", "l", "r", "r", "l"),
  caption = paste0(
    "All ", n_counties, " Great Britain counties ordered by decreasing number of neighbours, ",
    "ties broken by decreasing shared boundary length (km). ",
    "Computed in EPSG:27700 (British National Grid). ",
    "Source: ", gadm_link, " polygons; adjacency via ", sf_link, "."
  )
)
```
